---
title: "Machine Learnign Project"
output: html_notebook
---

This assignment is based on the work done by Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013

Read more: http://groupware.les.inf.puc-rio.br/har#ixzz5KiZnE8Ua 
```{r message = FALSE, warning=FALSE}
suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(mboost))
suppressPackageStartupMessages(library(dplyr))
```

```{r}
training <-as_tibble(read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"))
testing <-as_tibble(read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"))

```

Looking at the data it seems like a lot of the rows have NAs.  Let's see how they are distributed.
```{r}
percentNa <- function(x){ sum(is.na(x))/length(x)}
table(apply(training, 2, percentNa))
```
Let's just drop the columns that are mostly NAs, and do a 10 fold cross validation on a random forest.

```{r}
training[, c(apply(training,2,percentNa)< 0.5)] -> reducedtraining
```
Still too many features, could not get analysis to complete.  Maybe I can try and reduce the dimensionality a little more, and look at some of the subsets.  There are groups of data named belt, arm, forearm, dumbell.  Lets see if any of those are interesting.
```{r}
reducedtraining %>% select(contains("_belt")) %>% select_if(is.numeric) %>% mutate(classe = reducedtraining$classe) -> beltdata
reducedtraining %>% select(contains("_arm")) %>% select_if(is.numeric) %>% mutate(classe = reducedtraining$classe) -> armdata
reducedtraining %>% select(contains("_fore")) %>% select_if(is.numeric) %>% mutate(classe = reducedtraining$classe) -> forearmdata
reducedtraining %>% select(contains("_dumb")) %>% select_if(is.numeric) %>% mutate(classe = reducedtraining$classe) -> dumbbelldata
```
We now have 4 sets of data with around 20 features each.
```{r}
#featurePlot(armdata[,-14], armdata$classe, plot = "pairs")
#featurePlot(beltdata[,-14], beltdata$classe, plot = "pairs")
#featurePlot(dumbbelldata[,-14], dumbbelldata$classe, plot = "pairs")
#featurePlot(forearmdata[,-14], forearmdata$classe, plot = "pairs")
```
Armdata(1,2,8,11,12,13) look like they have some separation.
Beltdata(7,12) look like they have some separation.
Dumbbelldata(4-8,10-13) have separation.
Forearmdata(2,4-8) have separation

```{r}

interesting_features <- c("roll_arm","pitch_arm","accel_arm_x","magnet_arm_x", "magnet_arm_y", "magnet_arm_z", "gyros_belt_z","magnet_belt_y","total_accel_dumbbell","gyros_dumbbell_x", "gyros_dumbbell_y", "gyros_dumbbell_z","accel_dumbbell_x","accel_dumbbell_z","magnet_dumbbell_x","magnet_dumbbell_y", "magnet_dumbbell_z", "pitch_forearm", "total_accel_forearm", "gyros_forearm_x", "gyros_forearm_y", "gyros_forearm_z" )
```

```{r}
#Make parrallel
library(parallel)
library(doParallel)
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)



# define training control
#train_control<- trainControl(method="cv", number = 3, allowParallel = TRUE, seeds = list(c(1:3),c(4:6),c(7:9),10 ))  #.9798 accuracy
train_control<- trainControl(method="cv", number = 5, allowParallel = TRUE, seeds = list(c(1:3),c(4:6),c(7:9), c(10:12), c(13:15), 16 ))
features = paste(interesting_features, collapse = "+")
training %>% select(one_of(interesting_features)) %>% mutate (classe = training$classe) -> smallSet
#set.seed(42)
#inTrain <- createDataPartition(y=smallSet$classe, times=1, p=0.1, list=FALSE)
#smalltrain <- reducedtraining[inTrain,]
model<- train(classe ~ ., data=smallSet,  method="rf", trControl = train_control)

#Stop parallel processing
stopCluster(cluster)
registerDoSEQ()

print(model)
```
```{r}
print nearZeroVar table
nearZeroVar(training,saveMetrics=TRUE)
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

